---
title: 'Puerto Rico Remote Sensing - Belowground '
output:
  html_document:
    code_folding: hide
    df_print: paged
  pdf_document: default
---
# Resin P Analysis - SB2 & EV1
Pre-hurricane analysis of resin P on sites SB2 and EV1. Collected and processed March 2017.
```{r Dataset}
library(tidyverse)
resinp <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/ResinP.csv", header = TRUE, stringsAsFactors = TRUE, na.strings = "NA") # units = ugP/g soil dry weight
```

## Resin P Summary Stats
```{r}
# resin p mean, sd, and se by site
resinp %>% 
  group_by(Site) %>% 
  summarise(
     n = n(),
    mean = mean(ResinP, na.rm = TRUE),
    sd_total = sd(ResinP, na.rm = TRUE),
    se_total = (sd(ResinP, na.rm = TRUE)/(sqrt(n()))))

# resin p mean, sd, and se by site x tree species
resinp %>% 
  group_by(Site, TreeSpecies) %>% 
  summarise(
     n = n(),
    mean = mean(ResinP, na.rm = TRUE),
    sd_total = sd(ResinP, na.rm = TRUE),
    se_total = (sd(ResinP, na.rm = TRUE)/(sqrt(n()))))
```

## Resin P ANOVA
```{r}
resinp.aov <- aov(ResinP ~ Site + TreeSpecies + Site*TreeSpecies, data = resinp)
summary(resinp.aov)

TukeyHSD(resinp.aov)
```

Resin P differed between EV1 and SB2 (p-value < 0.05), driven only by site. Tukey HSD shows that there were no significant differences between tree species within and among sites. 

## Resin P Graphs
### Resin P by Site
```{r}
# vector for tick marks
ticks <- seq(0, 2, by = 0.20)

ggplot(data = resinp, aes(x = Site, y = ResinP)) + geom_boxplot() + geom_point(aes(x = Site, y = ResinP, colour = TreeSpecies), position = position_dodge(0.1)) + scale_y_continuous(limits = c(0,2), breaks = ticks) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Resin~P~(mu~g~P~g^{-1}))))
```

### Resin P by Site & Tree Species
```{r}
ggplot(data = resinp, aes(x = Site, y = ResinP, colour = TreeSpecies), position_dodge(position_dodge(0.5))) + geom_boxplot() + geom_point(aes(x = Site, y = ResinP, colour = TreeSpecies), position = position_dodge(0.5)) + scale_y_continuous(limits = c(0,2), breaks = ticks) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Resin~P~(mu~g~P~g^{-1}))))
```

# Organic Phosphorus
Organic phosphorus results from SB2 and EV1 from February 2018 Puerto Rico sampling trip. 

```{r Organic phosphorus}

soilp <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/SoilPhosphorus.csv", header = TRUE, na.strings = "NA", stringsAsFactors = TRUE)
```

## Acid organic phosphorus
```{r Acid organic phosphorus}
ggplot(data = soilp, aes(x = Site, y = Acid_OP)) + geom_boxplot() + geom_point(aes(x = Site, y = Acid_OP, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Acid~Organic~P~(mg~P~L^{-1}))))
```

## Alkaline organic phosphorus
```{r Alkaline organic phosphorus}
ggplot(data = soilp, aes(x = Site, y = Alkaline_OP)) + geom_boxplot() + geom_point(aes(x = Site, y = Alkaline_OP, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Alkaline~Organic~P~(mg~P~L^{-1}))))
```

## Total organic phosphorus
```{r Total organic phosphorus}
ggplot(data = soilp, aes(x = Site, y = Total_OP)) + geom_boxplot() + geom_point(aes(x = Site, y = Total_OP, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Total~Organic~P~(mg~P~L^{-1}))))

soilp.siteaov <- aov(Total_OP ~ Site, data = soilp)
summary(soilp.siteaov)
```

## Average organic phosphorus at each site
```{r Organic phosphorus by site}
soilp.sum <- soilp %>% 
  group_by(Site) %>% 
  summarise(
    n = n(),
    mean_total = mean(Total_OP, na.rm = TRUE),
    sd_total = sd(Total_OP, na.rm = TRUE),
    se_total = (sd(Total_OP, na.rm = TRUE)/(sqrt(n()))))
View(soilp.sum)

ggplot(data = soilp.sum, aes(x = Site, y = mean_total)) + geom_bar(stat = "identity", colour = "black") + geom_errorbar(aes(ymax=mean_total+sd_total, ymin=mean_total-sd_total), width = 0.2, size = 0.3) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Mean~Organic~P~(mg~P~L^{-1})))) 

```

Total organic phosphorus is the sum of acid and alkaline extracted phosphorus. Acid extracted organic phosphorus comprised the majority of the total organic phosphorus at both EV1 and SB2, with significantly (p-value < 0.001) higher amounts of organic P at SB2 (mean = 7.85 mgP/L) relative to EV1 (mean = 4.98 mgP/L). 

## Organic phosphorus by site and tree species
```{r Organic phosphorus by site and tree species}
soilp.sum2 <- soilp %>% 
  group_by(Site, TreeSpecies) %>% 
  summarise(
    n = n(),
    mean_total = mean(Total_OP, na.rm = TRUE),
    sd_total = sd(Total_OP, na.rm = TRUE),
    se_total = (sd(Total_OP, na.rm = TRUE)/(sqrt(n()))))
View(soilp.sum2)

ggplot(data = soilp.sum2, aes(x = Site, y = mean_total, colour = TreeSpecies)) + geom_point(position = position_dodge(0.5)) + geom_errorbar(aes(ymax=mean_total+se_total, ymin=mean_total-se_total), position = position_dodge(0.5)) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Mean~Organic~P~(mg~P~L^{-1})))) 

soilp.sitespecies.aov <- aov(Total_OP ~ Site * TreeSpecies, data = soilp)
summary(soilp.sitespecies.aov)

TukeyHSD(soilp.sitespecies.aov)
```

When tree species is taken into account, site is still the main driver of organic phosphorus availability (p-value < 0.001) though tree species may play a slight role (p-value < 0.1). 

# Root phosphatase activity
```{r SB2 & EV1 root phosphatase}
pase <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/RemoteSensing_Pase.csv", header = TRUE, stringsAsFactors = TRUE, na.strings = "NA")
```

## Root phosphatase by site
```{r Root pase x site}
pase %>%
  group_by(Site) %>% 
  summarise(
     n = n(),
    mean = mean(Pase, na.rm = TRUE),
    sd_total = sd(Pase, na.rm = TRUE),
    se_total = (sd(Pase, na.rm = TRUE)/(sqrt(n()))))
```

```{r Root pase x site graph}
ggplot(data = pase, aes(x = Site, y = Pase)) + geom_boxplot() + geom_point(aes(x = Site, y = Pase, colour = TreeSpecies)) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + xlab("Tree Species") + ylab(expression(Root~phosphatase~(mu~moles~pNP~g^{-1}))) + theme(legend.position = c(0.9,0.75))
```

## Root phosphatase by site & tree species
```{r Root pase x site x tree species}
pase %>%
  group_by(Site, TreeSpecies) %>% 
  summarise(
     n = n(),
    mean = mean(Pase, na.rm = TRUE),
    sd_total = sd(Pase, na.rm = TRUE),
    se_total = (sd(Pase, na.rm = TRUE)/(sqrt(n()))))
```

```{r Root pase x site x tree species plot}
ggplot(data = pase, aes(x = reorder(TreeSpecies, Pase, FUN = median), y = Pase, fill = TreeSpecies)) + geom_boxplot() + geom_point(aes(x = TreeSpecies, y = Pase, colour = TreeSpecies)) + facet_grid(. ~ Site) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + xlab("Tree Species") + ylab(expression(Root~phosphatase~(mu~moles~pNP~g^{-1}))) + theme(legend.position = c(0.1,0.8))
```

## Root phoosphatase ANOVA
```{r Root pase ANOVA}
rootpase.aov <- aov(Pase ~ Site + TreeSpecies + Site*TreeSpecies, data = pase)
summary(rootpase.aov)  

TukeyHSD(rootpase.aov)
```

Root phosphatase is influenced by both Site and Tree Species (p-value < 0.001), though independently. In genenral, PREMON differs from CECSCH and INGLAU across both sites (p-value <0.05). As expected, PREMON-INGLAU differences are greatest when at different sites (p-value < 0.01) than when PREMON nad INGLAU are at the same site (p-value <0.05). 