---
title: "Puerto Rico Remote Sensing - Belowground "
output:
  html_document:
    df_print: paged
    code_folding: hide
---
# Resin P Analysis
Resin P results from ICA, EVR, EVV, SB2 and EV1 from all previous trips to Puerto Rico.


```{r Resin P}
library(tidyverse)
allresin <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/ResinP_5Sites.csv")

allresin %>%
  group_by(Site) %>% 
  summarise(mean = mean(ugP.L))

allresin$Site <- reorder(allresin$Site, allresin$ugP.L)
ggplot(data = allresin, aes(x = Site, y = ugP.L)) + geom_boxplot() + theme_bw(base_size = 20) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(axis.title.x = element_blank()) + ylab((expression(Resin~~P~(mu~g~P~L^{-1}))))

resin.aov <- aov(ugP.L ~ Site, data = allresin)
summary(resin.aov)

TukeyHSD(resin.aov)
```

Resin P differs across experimental sites (p-value < 0.001) with Rio Icacos (ICA) containing the least available amounts of orthophosphate (mean = 19.66 ugP/L) and Sabana 2 (SB2) with the highest (mean = 98.00 ugP/L). Tukey HSD results indicate that the differences across sites were most significant (p-value < 0.001) between SB2-ICA, SB2-EV1, and SB2-El Verde Ridge (EVR). 

# Organic Phosphorus
Organic phosphorus results from SB2 and EV1 from February 2018 Puerto Rico sampling trip. 

```{r Organic phosphorus}

soilp <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/SoilPhosphorus.csv", header = TRUE, na.strings = "NA", stringsAsFactors = TRUE)
```

## Acid organic phosphorus
```{r Acid organic phosphorus}
ggplot(data = soilp, aes(x = Site, y = Acid_OP)) + geom_boxplot() + geom_point(aes(x = Site, y = Acid_OP, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Acid~Organic~P~(mg~P~L^{-1}))))
```

## Alkaline organic phosphorus
```{r Alkaline organic phosphorus}
ggplot(data = soilp, aes(x = Site, y = Alkaline_OP)) + geom_boxplot() + geom_point(aes(x = Site, y = Alkaline_OP, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Alkaline~Organic~P~(mg~P~L^{-1}))))
```

## Total organic phosphorus
```{r Total organic phosphorus}
ggplot(data = soilp, aes(x = Site, y = Total_OP)) + geom_boxplot() + geom_point(aes(x = Site, y = Total_OP, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Total~Organic~P~(mg~P~L^{-1}))))

soilp.siteaov <- aov(Total_OP ~ Site, data = soilp)
summary(soilp.siteaov)
```

## Average organic phosphorus at each site
```{r Organic phosphorus by site}
soilp.sum <- soilp %>% 
  group_by(Site) %>% 
  summarise(
    n = n(),
    mean_total = mean(Total_OP, na.rm = TRUE),
    sd_total = sd(Total_OP, na.rm = TRUE),
    se_total = (sd(Total_OP, na.rm = TRUE)/(sqrt(n()))))
View(soilp.sum)

ggplot(data = soilp.sum, aes(x = Site, y = mean_total)) + geom_bar(stat = "identity", colour = "black") + geom_errorbar(aes(ymax=mean_total+sd_total, ymin=mean_total-sd_total), width = 0.2, size = 0.3) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Mean~Organic~P~(mg~P~L^{-1})))) 

```

Total organic phosphorus is the sum of acid and alkaline extracted phosphorus. Acid extracted organic phosphorus comprised the majority of the total organic phosphorus at both EV1 and SB2, with significantly (p-value < 0.001) higher amounts of organic P at SB2 (mean = 7.85 mgP/L) relative to EV1 (mean = 4.98 mgP/L). 

## Organic phosphorus by site and tree species
```{r Organic phosphorus by site and tree species}
soilp.sum2 <- soilp %>% 
  group_by(Site, TreeSpecies) %>% 
  summarise(
    n = n(),
    mean_total = mean(Total_OP, na.rm = TRUE),
    sd_total = sd(Total_OP, na.rm = TRUE),
    se_total = (sd(Total_OP, na.rm = TRUE)/(sqrt(n()))))
View(soilp.sum2)

ggplot(data = soilp.sum2, aes(x = Site, y = mean_total, colour = TreeSpecies)) + geom_point(position = position_dodge(0.5)) + geom_errorbar(aes(ymax=mean_total+se_total, ymin=mean_total-se_total), position = position_dodge(0.5)) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Mean~Organic~P~(mg~P~L^{-1})))) 

soilp.sitespecies.aov <- aov(Total_OP ~ Site * TreeSpecies, data = soilp)
summary(soilp.sitespecies.aov)

TukeyHSD(soilp.sitespecies.aov)
```

When tree species is taken into account, site is still the main driver of organic phosphorus availability (p-value < 0.001) though tree species may play a slight role (p-value < 0.1). 
