---
title: 'Puerto Rico Remote Sensing - Belowground '
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float: true
  
---
# Resin P Analysis
Pre-hurricane analysis of resin P on sites SB2 and EV1. Collected and processed March 2017.

```{r Dataset}
library(tidyverse)
resinp <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/ResinP_2017.csv", header = TRUE, stringsAsFactors = TRUE, na.strings = "NA") # units = ugP/g soil dry weight
```

## Resin P Summary Stats

### Resin P at SB2 and EV1
```{r Resin P x Site}
# resin p mean, sd, and se by site
resinp %>% 
  group_by(Site) %>% 
  summarise(
     n = n(),
    mean = mean(ResinP, na.rm = TRUE),
    sd = sd(ResinP, na.rm = TRUE),
    se = (sd(ResinP, na.rm = TRUE)/(sqrt(n()))))

# graph
ticks <- seq(0, 2, by = 0.20)

ggplot(data = resinp, aes(x = Site, y = ResinP)) + geom_boxplot() + geom_point(aes(x = Site, y = ResinP, colour = TreeSpecies), position = position_dodge(0.1)) + scale_y_continuous(limits = c(0,2), breaks = ticks) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Resin~P~(mu~g~P~g^{-1}))))
```

### Resin P near tree species' at SB2 and EV1
```{r Resin P x Site, Species}
# resin p mean, sd, and se by site x tree species
resinp %>% 
  group_by(Site, TreeSpecies) %>% 
  summarise(
     n = n(),
    mean = mean(ResinP, na.rm = TRUE),
    sd_total = sd(ResinP, na.rm = TRUE),
    se_total = (sd(ResinP, na.rm = TRUE)/(sqrt(n()))))

# graph
ggplot(data = resinp, aes(x = Site, y = ResinP, colour = TreeSpecies), position_dodge(position_dodge(0.5))) + geom_boxplot() + geom_point(aes(x = Site, y = ResinP, colour = TreeSpecies), position = position_dodge(0.5)) + scale_y_continuous(limits = c(0,2), breaks = ticks) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Resin~P~(mu~g~P~g^{-1}))))
```

Resin P is highest at SB2 (0.874 +/- 0.758) relative to EV1 (0.139 +/- 0.217), consistent with the highest resin P averages near CALCAL and PREMON trees in SB2. When comparing the same tree species across the two sites, resin p values near trees at SB2 are consistently higher. 

## Resin P ANOVA
```{r ANOVA}
resinp.aov <- aov(ResinP ~ Site + TreeSpecies + Site*TreeSpecies, data = resinp)
summary(resinp.aov)

TukeyHSD(resinp.aov)
```

As expected, resin P differed significantly between EV1 and SB2 (p-value < 0.05), driven only by site. Tukey HSD shows that there were no significant differences between tree species within and among sites.

# Organic Phosphorus
Organic phosphorus results from SB2 and EV1 from March 2017 Puerto Rico sampling trip. 

```{r Organic phosphorus data}
organicp <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/RemoteSensing_OrganicP.csv", header = TRUE, na.strings = "NA", stringsAsFactors = TRUE)
```

## Acid organic phosphorus
```{r Acid OP Plot}
ggplot(data = organicp, aes(x = Site, y = Acid)) + geom_boxplot() + geom_point(aes(x = Site, y = Acid, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Acid~Organic~P~(mu~g~P~g^{-1}))))
```

## Alkaline organic phosphorus
```{r Alkaline OP Plot}
ggplot(data = organicp, aes(x = Site, y = Alkaline)) + geom_boxplot() + geom_point(aes(x = Site, y = Alkaline, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Alkaline~Organic~P~(mu~g~P~g^{-1}))))
```

## Total organic phosphorus
```{r Total OP Plot}
ggplot(data = organicp, aes(x = Site, y = Total)) + geom_boxplot() + geom_point(aes(x = Site, y = Total, colour = TreeSpecies), position = position_dodge(0.1), size = 3.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Total~Organic~P~(mu~g~P~g^{-1}))))

organicp.siteaov <- aov(Total ~ Site, data = organicp)
summary(organicp.siteaov)
```
Total organic phosphorus is the sum of acid and alkaline extracted phosphorus. Acid extracted organic phosphorus comprised the majority of the total organic phosphorus at both EV1 and SB2, with significantly (p-value < 0.001) higher amounts of organic P at SB2 (mean = 7.85 mgP/L) relative to EV1 (mean = 4.98 mgP/L). 

## Average organic phosphorus at each site
```{r Organic phosphorus x site}
organicp.sum <- organicp %>% 
  group_by(Site) %>% 
  summarise(
    n = n(),
    mean_total = mean(Total, na.rm = TRUE),
    sd_total = sd(Total, na.rm = TRUE),
    se_total = (sd(Total, na.rm = TRUE)/(sqrt(n()))))
View(organicp.sum)

ggplot(data = organicp.sum, aes(x = Site, y = mean_total)) + geom_bar(stat = "identity", colour = "black") + geom_errorbar(aes(ymax=mean_total+sd_total, ymin=mean_total-sd_total), width = 0.2, size = 0.3) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Mean~Organic~P~(mu~g~P~g^{-1})))) 

```

## Organic phosphorus by site and tree species
```{r Organic phosphorus x site and tree species}
organicp.sum2 <- organicp %>% 
  group_by(Site, TreeSpecies) %>% 
  summarise(
    n = n(),
    mean_total = mean(Total, na.rm = TRUE),
    sd_total = sd(Total, na.rm = TRUE),
    se_total = (sd(Total, na.rm = TRUE)/(sqrt(n()))))
View(organicp.sum2)

ggplot(data = organicp.sum2, aes(x = Site, y = mean_total, colour = TreeSpecies)) + geom_point(position = position_dodge(0.5)) + geom_errorbar(aes(ymax=mean_total+se_total, ymin=mean_total-se_total), position = position_dodge(0.5)) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  legend.position = c(0.1,0.8)) + ylab((expression(Mean~Organic~P~(mu~g~P~g^{-1})))) 

organicp.sitespecies.aov <- aov(Total ~ Site * TreeSpecies, data = organicp)
summary(organicp.sitespecies.aov)

TukeyHSD(organicp.sitespecies.aov)
```

When tree species is taken into account, site is still the main driver of organic phosphorus availability (p-value < 0.001) though tree species may play a slight role (p-value < 0.1). 

# Root phosphatase activity
```{r Root phosphatase data}
pase <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/RemoteSensing_Pase.csv", header = TRUE, stringsAsFactors = TRUE, na.strings = "NA")
```

## Root phosphatase by site
```{r Root pase x site}
pase %>%
  group_by(Site) %>% 
  summarise(
     n = n(),
    mean = mean(Pase, na.rm = TRUE),
    sd_total = sd(Pase, na.rm = TRUE),
    se_total = (sd(Pase, na.rm = TRUE)/(sqrt(n()))))

# graph
ggplot(data = pase, aes(x = Site, y = Pase)) + geom_boxplot() + geom_point(aes(x = Site, y = Pase, colour = TreeSpecies)) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + xlab("Tree Species") + ylab(expression(Root~phosphatase~(mu~moles~pNP~g^{-1}))) + theme(legend.position = c(0.9,0.75))
```

## Root phosphatase by site & tree species
```{r Root pase x site and tree species}
pase %>%
  group_by(Site, TreeSpecies) %>% 
  summarise(
     n = n(),
    mean = mean(Pase, na.rm = TRUE),
    sd_total = sd(Pase, na.rm = TRUE),
    se_total = (sd(Pase, na.rm = TRUE)/(sqrt(n()))))

# graph
ggplot(data = pase, aes(x = reorder(TreeSpecies, Pase, FUN = median), y = Pase, fill = TreeSpecies)) + geom_boxplot() + geom_point(aes(x = TreeSpecies, y = Pase, colour = TreeSpecies)) + facet_grid(. ~ Site) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + xlab("Tree Species") + ylab(expression(Root~phosphatase~(mu~moles~pNP~g^{-1}))) + theme(legend.position = c(0.1,0.8))
```

## Root phoosphatase ANOVA
```{r Root pase ANOVA}
rootpase.aov <- aov(Pase ~ Site + TreeSpecies + Site*TreeSpecies, data = pase)
summary(rootpase.aov)  

TukeyHSD(rootpase.aov)
```

Root phosphatase is influenced by both Site and Tree Species (p-value < 0.001), though independently. In genenral, PREMON differs from CECSCH and INGLAU across both sites (p-value <0.05). As expected, PREMON-INGLAU differences are greatest when at different sites (p-value < 0.01) than when PREMON nad INGLAU are at the same site (p-value <0.05). 

# Root phosphatase and soil phosphorus
```{r Belowground data 2017}
bg.2017 <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/PR2_OrganicResinPase.csv", header = TRUE, stringsAsFactors = TRUE, na.strings = "NA")
```

```{r Site Summary}
resinp.orgp.sum2 <- bg.2017 %>% 
  group_by(Site) %>% 
  summarise(
    n = n(),
    resin_mean = mean(ResinP, na.rm = TRUE),
    resin_sd = sd(ResinP, na.rm = TRUE),
    resin_se = (sd(ResinP, na.rm = TRUE)/(sqrt(n()))),
    orgP_mean = mean(Total_OrganicP, na.rm = TRUE),
    orgP_sd = sd(Total_OrganicP, na.rm = TRUE),
    orgP_se = (sd(Total_OrganicP, na.rm = TRUE)/(sqrt(n()))),
    pase_mean = mean(Phosphatase, na.rm = TRUE),
    pase_sd = sd(Phosphatase, na.rm = TRUE),
    pase_se = (sd(Phosphatase, na.rm = TRUE)/(sqrt(n()))))
View(resinp.orgp.sum2)
```

## Resin P by Site
```{r}
ggplot(data = resinp.orgp.sum2, aes(x = Site, y = resin_mean)) + geom_bar(stat = "identity", colour = "black") + geom_errorbar(aes(ymax=resin_mean+resin_se, ymin=resin_mean-resin_se), width = 0.2, size = 0.3) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank()) + ylab((expression(Mean~Resin~P~(mu~g~P~g^{-1})))) 
```

## Organic P by Site
```{r}
ggplot(data = resinp.orgp.sum2, aes(x = Site, y = orgP_mean)) + geom_bar(stat = "identity", colour = "black") + geom_errorbar(aes(ymax=orgP_mean+orgP_se, ymin=orgP_mean-orgP_se), width = 0.2, size = 0.3) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank()) + ylab((expression(Mean~Organic~P~(mu~g~P~g^{-1}))))
```

## Root phosphatase by Site
```{r}
ggplot(data = resinp.orgp.sum2, aes(x = Site, y = pase_mean)) + geom_bar(stat = "identity", colour = "black") + geom_errorbar(aes(ymax=pase_mean+pase_se, ymin=pase_mean-pase_se), width = 0.2, size = 0.3)  + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank()) + ylab((expression(Mean~Root~Phosphatase~(mu~moles~P~g^{-1}))))
```

## Comparisons

### Resin P and Organic P
```{r}
ggplot(data =  bg.2017, aes(x = Total_OrganicP, y = ResinP, colour = Tree.Species, shape = Site)) + geom_point(size = 5.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank()) + ylab((expression(Resin~P~({mu}~g~P~g^{-1})))) +  xlab((expression(Organic~P~(mu~g~P~g^{-1}))))
```

### Ratio of Resin P:Organic P
```{r}
bg.2017 <- bg.2017 %>% 
  mutate(Ratio_ResinOrganic = ResinP / Total_OrganicP)

ggplot(data = bg.2017, aes(x = Site, y = Ratio_ResinOrganic, colour = Tree.Species)) + geom_point(position = position_dodge(0.5)) + scale_y_continuous(limits = c(0, 5)) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank()) + ylab("ResinP : Organic P") + theme(legend.position = c(0.1, 0.8))
  
```

### Root phosphatase x Resin P
```{r}
ggplot(data =  bg.2017, aes(x = ResinP, y = Phosphatase, colour = Tree.Species, shape = Site)) + geom_point(size = 5.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank()) + ylab((expression(Root~Phosphatase~({mu}~moles~P~g^{-1})))) +  xlab((expression(Resin~P~(mu~g~P~g^{-1}))))
```

### Root phosphatase x Organic P
```{r}
ggplot(data =  bg.2017, aes(x = Total_OrganicP, y = Phosphatase, colour = Tree.Species, shape = Site)) + geom_point(size = 5.0) + theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank()) + ylab((expression(Root~Phosphatase~({mu}~moles~P~g^{-1})))) +  xlab((expression(Organic~P~(mu~g~P~g^{-1}))))
```

# Microbial Analysis
```{r Microbial Dataset}
microbial <- read.csv("~/Documents/LabStats/PR2_RemoteSensing/Microbial_SoilPase.csv", header = TRUE, stringsAsFactors = TRUE, na.strings = "NA")
```

## Summary Statistics
```{r}
library(plotrix)
microbial.sum <- microbial %>%
  group_by(Site, TreeSpecies) %>% 
  summarise_at(c("TotalCarbon", "TotalNitrogen", "TC.TN", "NH4", "NO3", "PO4", "ExtC", "ExtN", "microbialC", "microbialN", "microbialP", "microbialCN", "TotalOrganicP", "ResinP", "Rootpase"), funs(mean, sd, std.error), na.rm = TRUE)
View(microbial.sum)

# Total Carbon
ggplot(data = microbial.sum, aes(x = Site, y = TotalCarbon_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = TotalCarbon_mean))

# Total Nitrogen 
ggplot(data = microbial.sum, aes(x = Site, y = TotalNitrogen_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = TotalNitrogen_mean)) # see if root pase is higher with higher total N

# TC/TN mean
ggplot(data = microbial.sum, aes(x = Site, y = TC.TN_mean)) + geom_boxplot()

# NH4 mean
ggplot(data = microbial.sum, aes(x = Site, y = NH4_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = NH4_mean))

# NO3 mean
ggplot(data = microbial.sum, aes(x = Site, y = NO3_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = NO3_mean))

# PO4 mean
ggplot(data = microbial.sum, aes(x = Site, y = PO4_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = PO4_mean)) # consistent with resin P

# ExtC
ggplot(data = microbial.sum, aes(x = Site, y = ExtC_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = ExtC_mean))

# ExtN
ggplot(data = microbial.sum, aes(x = Site, y = ExtN_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = ExtN_mean))

# microbialC
ggplot(data = microbial.sum, aes(x = Site, y = microbialC_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = microbialC_mean)) # consistent with lower amounts of organic phosphorus at each site

# microbial N
ggplot(data = microbial.sum, aes(x = Site, y = microbialN_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = microbialN_mean))

# microbial P
ggplot(data = microbial.sum, aes(x = Site, y = microbialP_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = microbialP_mean))

# microbial CN
ggplot(data = microbial.sum, aes(x = Site, y = microbialCN_mean)) + geom_boxplot() + geom_point(aes(x = Site, y = microbialCN_mean))
```

## ANOVAs
```{r}
# Total Carbon
tc.aov <- aov(TotalCarbon ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(tc.aov) # Site

# Total Nitrogen
tn.aov <- aov(TotalNitrogen ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(tn.aov) # none

# TC/TN
tctn.aov <- aov(TC.TN ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(tctn.aov)

TukeyHSD(tctn.aov)

# Nitrogen and Phosphorus
nh4.aov <- aov(NH4 ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(nh4.aov) # none

no3.aov <- aov(NO3 ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(no3.aov) # significant

TukeyHSD(no3.aov)

po4.aov <- aov(PO4 ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(po4.aov) # Site

# Extractable C and N
extc.aov <- aov(ExtC ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(extc.aov) # site

extn.aov <- aov(ExtN ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(extn.aov) # tree species

# Microbial 
microC.aov <- aov(microbialC ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(microC.aov) # Site

microN.aov <- aov(microbialN ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(microN.aov) # Site

microP.aov <- aov(microbialP ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(microP.aov) # Site

microCN.aov <- aov(microbialCN ~ Site + TreeSpecies + Site:TreeSpecies, data = microbial)
summary(microCN.aov)
```

Total carbon differs between EV1 and SB2 significantly (p< 0.05), with EV1 containing higher total carbon. Total nitrogen did not differ between site or tree species. TC/TN ratios were highly significant (p<0.01), with EV1 having higher TC/TN ratios. NO3 concentrations are significantly lower in EV1 than in SB2 with significant differences between tree species from each site. Similar to our results obtained using resin P, Bray 1 showed that PO4 was significantly (p <0.5) lower in EV1 than in SB2. Although extractable C was only significantly different between sites, extractable N was only significantly different between tree species(p<0.05), though with an interaction term (p<0.05). Microbial C,N, and P are significantly lower in EV1 than in SB2 (p<0.05), though microbial CN is not significantly different at both sites. (edit later for wording) 

## Correlation Matrix
```{r}
library(corrplot)

micro2 <- microbial[-c(3), ] # removing data with resin P NA
microbial.cor <- cor(micro2[, c(4:18)]) # specifying which rows to create matrix from
M <- corrplot(microbial.cor, method = "color") 

res1 <- cor.mtest(microbial.cor, conf.level = 0.95)
res2 <- cor.mtest(microbial.cor, conf.level = 0.99)

# plot
corrplot(M, p.mat = res1$p,insig = "label_sig", pch.col = "white", pch = "p<0.05", pch.cex = 0.5, order = "AOE")
```

